---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

display_msg <- function(checker) {
  attr(checker, "msg") |>
    str_c(collapse = "\n\n") |>
    cat()
}
```

# teal.builder.checks


# Overview

**teal.builder.checks** is a package for checking the data integrity of 
**ADaM** datasets

* [Package website] - (<https://github.com/teal.builder.checks/index.html>)

## Installation

You can install the development version from 
[Github](https://<github.com>/<teal.builder.checks>) with:

```{r, eval=FALSE, run=FALSE}

# remotes::install_github(
#  "https://<github.com>/teal.builder.checks",
#  auth_token = Sys.getenv("GITHUB_PAT"),
# upgrade = "never"
# )
```

# Usage

**teal.builder.checks** checks the following **ADaM** datasets:

| ADaM datasets |ADaM datasets contd. |   
|:-------------:|:-------------:|
| adam          | adrs          |
| adsl          | adqs          |
| adae          | adlb          |
| adtte         |               |

**teal.builder.checks** can perform the 
following data integrity checks by column on those datasets:

| Column Checks | Column Checks contd. | 
|:--------------|:--------------|
|existence | is date |
|is character | is datetime| 
|is integerish |has values |
|is double |lacks values |
|is logical |has one record per subject |
|is factor |has one record per subject per param |

# Example usage

```{r}
library(purrr)
library(dplyr)    
library(lubridate)
library(stringr)
library(checkmate)
library(vctrs)
library(random.cdisc.data)

# Load adam datasets for checking
adsl <- random.cdisc.data::radsl(N = 10, seed = 1, na_percentage = .1)
adtte <- random.cdisc.data::radtte(adsl, seed = 1, na_percentage = .1)
```
## Successful check of adsl

This check returns a logical **TRUE** with an attribute **data** which 
contains the original ADaM dataset checked


```{r}
library(teal.builder.checks)
check_adsl(adsl)
```

## Unsuccessful check of adsl

### check for non-existence of columns

This check returns a logical **FALSE** with two attributes

1) **msg** contains a character vector describing the reason for
failure, if the check failed

2) **data** which contains the ADaM dataset checked

```{r}
adsl |>
  dplyr::select(-STUDYID) |> 
  check_adsl()
```

## Creation of your own multichecker

### Successful check of adtte

In the following example, the checks are performed: 

1) the existence of certain columns 
2) specific values in one column
3) one record per subject per parameter

```{r}
my_adtte_checker <- make_multi_checker(
  check_has_cols(
    column_names = c("STUDYID", "USUBJID", "PARAMCD")
  ),
  check_has_values(
    column_name = "ARM",
    values = c("A: Drug X", "B: Placebo", "C: Combination")
  ),
  check_has_one_record_per_subject_per_param
)
```

```{r}
my_adtte_checker(adtte) 
```

## Unsuccessful check of adtte

This check returns a **msg** attribute with multiple error messages.

```{r}
adtte |>
  dplyr::slice_head(n = 12) |>
  dplyr::select(-STUDYID) |>
  dplyr::filter(ARM != "A: Drug X") |>
  my_adtte_checker() 
```
