---
title: "Checker function hierarchy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Checker function hierarchy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**teal.builder.checks** is a package for checking the data integrity of 
**ADaM** datasets


```{r setup}
library(teal.builder.checks)
```

# Overview

## Hierarchy of checker functions

**teal.builder.checks** checks the **ADaM** datasets 
and the functions are designed to be executed in a particular sequence.

The sequence starts with the most general checks 
i.e. Does the key column exist for a particular dataset?

then proceeds to data type checks such as **is_character** and **is_numeric**

and then finishes with the most specific checks 
such as **lacks_values**, **has one record per subject** 
and **has one record per subject per param**.

The execution sequence is as follows:

| Data Integrity Checks Sequence | Comments |
|:--------------|:--------------|
|existence | This should be done first to ensure later checks work |  
|is character | Datatype checks - the data type checks can be done in any order but should be done before the values checks and one record checks | 
|is integerish | Datatype checks - as above |
|is double | Datatype checks - as above |
|is logical | Datatype checks - as above |
|is factor | Datatype checks - as above | 
|is date | Datatype checks - as above |
|is datetime| Datatype checks - as above |
|has values | values checks - checking for certain values |
|lacks values | values checks - checking for the absence of certain values | 
|has one record per subject |one record check |
|has one record per subject per param | as above |

# Usage

### Hierarchy of functions example

An example of this is the **check_adrs** function:

```{r, eval=FALSE}
#' ADRS dataset checker
#'
 check_adrs <- function(adam) {
   col_vector <- c("STUDYID", "USUBJID")
   make_multi_checker(
     check_has_cols(col_vector),
     check_is_character("STUDYID"),
     check_is_character("USUBJID"),
     check_lacks_values("STUDYID", NA_character_),
     check_lacks_values("USUBJID", NA_character_),
     check_has_one_record_per_subject
 )(adam)
```

As can be seen from the code, 
first, it checks for the existence of columns,
then, it checks for character datatype in two columns,
then, it checks for **NA_character_** not being in the columns
and last, it checks that there is one record per subject.

Let's run the code:

First load the data:

```{r}
library(purrr)
library(dplyr)    
library(lubridate)
library(stringr)
library(checkmate)
library(vctrs)

# Load adam datasets for checking
adrs <- dplyr::tibble(
      "STUDYID" = c("AB12345", "AB12345", "AB12345"),
      "USUBJID" = c("AB12345-BRA-1-id-105", 
                    "AB12345-BRA-1-id-105",
                    "AB12345-BRA-1-id-134"),
      "SITEID" = c("BRA-1", "BRA-1", "BRA-1"),
      "PARAM" = c("Best Confirmed Overall Response by Investigator",
                  "Investigator End of Induction Response",
                  "Overall Response by Investigator"), 
      "PARAMCD" = c("BESRSPI", "INVET", "OVRINV"),
      "BOR" = c("PR", "SD", "SD"),
      "BORDSC" = c("Unconfirmed CR", 
                   "Unconfirmed CR", 
                   "Unconfirmed PR"
                   ),
      "BORUNC" = c("CR", "CR", "PR"),
      "RESPFL" = c("Y", "N", "N"),
      "DISCONFL" = c("Y", "Y", "Y"),
      "CLINBNFL" = c("Y", "Y", "Y"),
      "LPCHG" = c("Y", "Y", "Y"),
      "AVISIT" = c("Y", "Y", "Y")
    )

```


```{r}
library(teal.builder.checks)
check_adrs(adrs)
```
If we change the order of checking, then the error messages are given
but not in the most appropriate order:

```{r}
my_adrs_checker <- function(adam) {
   col_vector <- c("STUDYID", "USUBJID")
   make_multi_checker(
     check_has_one_record_per_subject, # move last checker to first 
     check_is_character("STUDYID"),
     check_is_character("USUBJID"),
     check_lacks_values("STUDYID", NA_character_),
     check_lacks_values("USUBJID", NA_character_),
     check_has_cols(col_vector) # move first checker to last
    
 )(adam)
}
```

As can seen in the output of **my_adrs_checker**, the most important message
is the last message 
**Dataset must have columns 'STUDYID', 'USUBJID', but is missing 'STUDYID'.** 
and the others result from that problem.

```{r}
adrs |> 
  dplyr::select(-STUDYID) |>
  my_adrs_checker() 
```

Whereas, the output of **check_adrs**, the most important message
is the only message 
**Dataset must have columns 'STUDYID', 'USUBJID', but is missing 'STUDYID'.** 
and the user has less output to check.

```{r}
adrs |> 
  dplyr::select(-STUDYID) |>
  check_adrs() 
```
