---
title: "Scope of datatype checker functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scope of datatype checker functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**teal.builder.checks** is a package for checking the data integrity of **ADaM** datasets

```{r setup}
library(teal.builder.checks)
```

# Overview - column existence check function


There is one checker function **check_has_cols** that checks for the existence of column/s.

# Usage - column existence check function

This function has one parameter **column_names** which can accept one column:

```{r}
 dplyr::tibble(STUDYID = "1") |>
   check_has_cols(column_names = "STUDYID")()
```

or a vector of columns

```{r}
dplyr::tibble(STUDYID = "1", USUBJID = "1") |>
   check_has_cols(column_names = c("STUDYID", "USUBJID"))()
```

# Overview - datatype checker functions

The datatype checkers check for particular datatype/s in columns.

The following table lists the datatype checkers and what they check:

| Datatype checker function | Description                             | Comments                                                 |
|:-----------------------|:-----------------------|:-----------------------|
| check_is_character        | checks if a column has character values | test done with base R **is.character**                   |
| check_is_integerish       | checks if a column has integer values   | test done with **checkmate::test_integerish**            |
| check_is_double           | checks if a column has numeric values   | test done with **vctrs::vec_is**                         |
| check_is_logical          | checks if a column has logical values   | test done with base R **is.logical**                     |
| check_is_categorical      | checks if a column has factor values    | test done with base R **is.factor** and **is.character** |
| check_is_date             | checks if a column has date values      | test done with **lubridate::is.Date**                    |
| check_is_datetime         | checks if a column has datetime values  | test done with **lubridate::is.POSIXct**                 |

# Usage - datatype checker functions

## **check_is_character**

This function returns **TRUE** when a column has a character datatype:

```{r}
comments <- "text"

dplyr::tibble(
   comments
 ) |> check_is_character("comments")()
```
This function returns **FALSE** and a message 
when a column does not have a character datatype:

```{r}
comments <- factor(c("male, female"))

dplyr::tibble(
   comments
 ) |> check_is_character("comments")()
```

## **check_is_integerish**

This function returns **TRUE** when a column has an integer datatype:

```{r}
USUBJID <- as.integer(1.1)

dplyr::tibble(
   USUBJID
) |> check_is_integerish("USUBJID")()

```

or a rounded double datatype:

```{r}
USUBJID <- as.double(1)

dplyr::tibble(
  USUBJID
) |> check_is_integerish("USUBJID")()
```

This function returns **FALSE** and a message 
when a column does not have an integer or rounded double datatype:

```{r}
USUBJID <- as.double(1.1)

dplyr::tibble(
  USUBJID
) |> check_is_integerish("USUBJID")()
```

## **check_is_double**

This function returns **TRUE** when a column has a double datatype:

```{r}
USUBJID <- as.double(1)

dplyr::tibble(
  USUBJID
) |> check_is_double("USUBJID")()
```

This function returns **FALSE** and a message 
when a column does not have a double datatype:

```{r}
test_datetime <- as.POSIXct("2020/01/01 22:22:22")

dplyr::tibble(
  test_datetime
) |> check_is_double("test_datetime")()
```

## **check_is_logical**

This function returns **TRUE** when a column has a logical datatype:

```{r}
status <- TRUE

dplyr::tibble(
  status
) |> check_is_logical("status")()
```

This function returns **FALSE** and a message 
when a column does not have a logical datatype:

```{r}
status <- factor(c("male, female"))

dplyr::tibble(
 status
) |> check_is_logical("status")()
```

## **check_is_categorical**

This function returns **TRUE** when a column has a categorical datatype:

```{r}
sex <- factor(c("male, female"))

dplyr::tibble(
  sex
) |> check_is_categorical("sex")()
```

or a character datatype:

```{r}
sex <- c("male, female")

dplyr::tibble(
  sex
) |> check_is_categorical("sex")()

```

This function returns **FALSE** and a message 
when a column does not have a categorical datatype:

```{r}
status <- as.double(1.1)

dplyr::tibble(
 status
) |> check_is_categorical("status")()
```

## **check_is_date**

This function returns **TRUE** when a column has a date datatype:

```{r}
library(lubridate)
test_date <- as.Date("2003-02-27")

dplyr::tibble(
  test_date
) |> check_is_date("test_date")()
```

This function returns **FALSE** and a message 
when a column does not have a date datatype:

```{r}
test_date <- as.POSIXct("2003-02-27 22:22:22")

dplyr::tibble(
  test_date
) |> check_is_date("test_date")()
```

## **check_is_datetime**

This function returns **TRUE** when a column has a **POSIXct** or **POSIXt** datatype:

```{r}
library(lubridate)
test_date <- as.POSIXct("2003-02-27 22:22:22")

dplyr::tibble(
  test_date
) |> check_is_datetime("test_date")()
```

This function returns **FALSE** and a message 
when a column does not have a **POSIXct** or **POSIXt** datatype:

```{r}
test_date <- as.Date("2003-02-27")

dplyr::tibble(
  test_date
) |> check_is_datetime("test_date")()
```
# Overview - values checker functions

There are two functions for checking for the presence 
or absence of particular values

| Datatype checker function | Description                             | Comments                                                 |
|:-----------------------|:-----------------------|:-----------------------|
| check_has_values        | checks if a column has specific values | good for testng for specific values                   |
| check_lacks_values        | checks if a column does not have specific values | good for testng for the lack of specific values e.g. **NA_character_**                   |
# Usage - values checker functions

## **check_has_values**

This function returns **TRUE** when a column has a specific value or values:

```{r}
dplyr::tibble(col_a = factor(letters)) |>
   check_has_values(column_name = "col_a", values = "a")()
```   

This function returns **FALSE** and a message 
when a column does not have a specific value or values:

```{r}
dplyr::tibble(col_a = factor(letters)) |>
   check_has_values(column_name = "col_a", values = c(1, 2))()
```   

## **check_lacks_values**

This function returns **TRUE** when a column does not have a specific value or values:

```{r}
dplyr::tibble(col_a = 1) |>
 check_lacks_values(column_name = "col_a", values = NA_character_)()
```   

This function returns **FALSE** and a message 
when a column does have a specific value or values:

```{r}
dplyr::tibble(col_a = NA_character_) |>
 check_lacks_values(column_name = "col_a", values = NA_character_)()
```   

# Usage - one record checker functions

## **check_has_one_record_per_subject**

This function returns **TRUE** when a dataset has one record per subject.
Thsi function is checking **USUBJID** against the records:

```{r}
dplyr::tibble(USUBJID = c(1, 2, 3), 
              col_a = c("a", "b", "c")) |>
    check_has_one_record_per_subject()
``` 

This function returns **FALSE** and a message 
when a dataset does not have one record per subject:

```{r}
dplyr::tibble(USUBJID = c(1, 1), col_a = c("a", "b")) |>
    check_has_one_record_per_subject()
```

## **check_has_one_record_per_subject_per_param**

This function returns **TRUE** when a dataset has one record per subject.
Thsi function is checking **USUBJID** against **PARAMCD** and the records:

```{r}
dplyr::tibble(
      STUDYID = c(1, 1, 1),
      USUBJID = c(1, 2, 3),
      PARAMCD = c(1, 1, 2)
    ) |>
    check_has_one_record_per_subject_per_param()
``` 

This function returns **FALSE** and a message 
when a dataset does not have one record per subject per param:

```{r}
dplyr::tibble(STUDYID = c(1, 1, 1), 
              USUBJID = c(1, 1, 2), 
              PARAMCD = c(1, 1, 2)) |>
    check_has_one_record_per_subject()
```
